World Heritage Sites on Wikipedia: Controversies in a Context of Constrained Agency for Cultural Heritage Activism

### 2020-05-31 analysis/paper/999-code-for-the-Stirling-ppt.Rmd
results section - talk pages
finished conclusion

# all WH sites, with and without WP pages (n = 846)
wh_wiki_table <-
  readxl::read_excel(here::here("analysis/data/raw_data/whc-sites-2019.xls"))

@@

wh_wiki_table_location <-
  wh_wiki_table %>%
  mutate(country = str_extract(Location,
                                 regex(country_names,
                                       ignore.case=TRUE))) %>%
    mutate(country = case_when(
      country == "United States" ~ "United States of America",
      country == "Czech Republic" ~ "Czechia",
      country == "Antigua & Barbuda" ~ "Antigua and Barb.",
      country == "Saint Kitts and Nevis" ~ "St. Kitts and Nevis",
      TRUE ~ as.character(country)))

>>>>

wh_wiki_table <-
  readxl::read_excel(here::here("analysis/data/raw_data/whc-sites-2019.xls"))

@@

wh_wiki_table_location <-
  wh_wiki_table %>%
  mutate(country = str_extract(states_name_en, # Location,
                                 regex(country_names,
                                       ignore.case=TRUE))) %>%
    mutate(country = case_when(
      country == "United States" ~ "United States of America",
      country == "Czech Republic" ~ "Czechia",
      country == "Antigua & Barbuda" ~ "Antigua and Barb.",
      country == "Saint Kitts and Nevis" ~ "St. Kitts and Nevis",
      TRUE ~ as.character(country)))

### 2020-05-31 analysis/paper/paper.Rmd
finished conclusion

---
title: "World Heritage Sites on Wikipedia: Inscription Controversies in a Context of Constrained Agency for Cultural Heritage Activism"
author:
  - Ben Marwick
  - Prema Smith
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  UNESCO World Heritage sites are places of outstanding significance, and often key sources of information that influence how people interact with the past today. The process of inscription on the UNESCO list is complicated and intersects with political and commercial controversies. But how well are these controversies known to the public? Wikipedia pages on these sites offer a unique dataset for insights into public understanding of heritage controversies. The unique technicity of Wikipedia, with its bot ecosystem and editing mechanics, shapes how knowledge about cultural heritage is constructed, and how controversies are negotiated and communicated. In this article we investigate the patterns of production, consumption and spatial and temporal distributions of Wikipedia pages for World Heritage cultural sites. We find that Wikipedia provides a distinctive context for investigating how people experience and relate to the past in the present. The agency of participants is highly constrained, but distinctive, behind-the-scenes expressions of cultural heritage activism are evident. Concerns about state-like actors, violence and destruction, deal-making, etc. in the World Heritage inscription process are poorly represented on Wikipedia’s World Heritage pages, instead hyper-local, and process issues dominate controversies on Wikipedia. We will consider how this kind of research, drawing on big data and data science methods, adds new dimensions to the debate on heritage controversies, especially its unique contributions, and also its limitations.
---

@@

An additional consideration for understanding participation in online peer production communities are the technical schemas of the software that Wikipedia runs on, MediaWiki. This is a complex toolkit that enables participation in Wikipedia in highly structured ways. On one hand, these structured behaviours produce structured datasets that are well suited to data science methods for efficient computational analysis of large numbers Wikipedia articles. On the other hand, they constrain and limit the agency of the user, canalizing their behaviour into a small number of possible actions and acceptable modes of discourse and engagement with other users [@iba2010analyzing]. While Wikipedia has elements that are ubiquitous on the Internet, such as links on articles that take the user to other articles or other pages on the Internet, it also has several less common elements that contribute to its unique technicity, resulting in very specific types of relationships between human users and the technical elements of the Wikipedia project [@niederer2010wisdom]. For example, every edit to an article is tracked in a publicly accessible version control system attached to that article. This exposes the article creation process in highly granular detail; for any given article, we can see how many editors contributed, the size of their edits and their distribution over time, among other things [@priedhorsky2007creating]. Wikipedia has a special category of edit called the 'revert' which allows a user to restore an article to an earlier state to remove recent vandalism (such as the addition of irrelevant or offensive material). This special revert action, combined with a 'talk' page attached to each encyclopedia article for threaded discussion among editors, allows us to detect and study the dynamics of conflicts arising from the creation and editing of articles [@suh2007us]. For example, a high proportion of 'revert' type edits, and a high word count on the talk page can indicate the controversiality of an article [@yasseri2012dynamics].

@@

In addition to the human users and the technological system that enables and constrains their activities on Wikipedia, there is an important third element of the ecosystem that contributes to Wikipedia's uniqueness: the bots. Wikipedia bots are computer scripts that automatically handle repetitive and mundane tasks to develop, improve, and maintain the encyclopedia. While bots are not unique to Wikipedia, they are important contributors, responsible for a large proportion of edits, and shape the human experience in significant ways [@niederer2010wisdom' @geiger2009social; @geiger2014bots]. They also evolve and autonomously engage in complex interactions with other bots to modify the encyclopedia [@tsvetkova2017even; @geiger2017operationalizing]. @zheng2019roles studied all 1,601 bots that were registered at the time of their study, and classified them into 9 bot roles and 25 associated bot functions. Table \@ref(tab:bot-table) shows the bots that were identified in our study and their categories and functions according to @zheng2019roles.

@@

We obtained data about Wikipedia articles by scraping the HTML pages with the rvest package for R [@rvest]. We used the SelectorGadget [@cantino2017selectorgadget] extension for the Chrome web browser to identify specific elements of interest, or nodes, on the HTML pages, and wrote custom R functions to extract data from these nodes. The highly structured nature of most Wikipedia pages afforded us much efficiency during data collection, however, a not insignificant number of pages had minor irregularities, for example, exotic table formatting or minor deviations from widely used templates, that necessitated more complex programming. An especially challenging issue was minor irregularities in spelling place names, for example, "Saint Kitts and Nevis" and "St. Kitts and Nevis", which can confound programmatic joining and summarizing of data by failing to recognize as the same place and leading to under-counting. We attempted to correct these manually in our code. Another challenge was in the formatting of tables, for example a merged cell in a table introduced an irregularity that made parsing and scraping complicated and fragile. We collected data during May 2019, and due to the dynamic nature of Wikipedia, it is highly likely that articles in our study have subtly changed since our data collection as users and bots continue their activity. It is unlikely that articles in our sample have been deleted from Wikipedia since our data collection, because of the high notability of CS-WHL, but it is possible that new articles have since appeared about sites that were not represented on Wikipedia since we collected the data. Our original scraping code may no longer work as the tables on Wikipedia articles continue to be modified.

@@

Only one of the controversial sites discussed above is in the group of articles where the talk page is longer than the article itself, and none of the site on the UNESCO heritage in danger list are in this group. The talk page for Mapungubwe is lengthy with expressions of concern about the accuracy of information in the article. In particular, editors were concerned about the depiction of Indigenous people, especially the degree to which early European colonists were aware of Indigenous communities, and the correct cultural affiliation of the Indigenous groups that originally occupied the site. Although not notable for the length of its talk page, the discussions among editors of the article about Preah Vihear Temple include a strongly worded disagreement among seven editors about the legitimacy of the ownership claims by Thailand and Cambodia of the territory that includes the temple. The debate among Wikipedia editors mirrors very closely the political tensions surrounding the nomination of Preah Vihear. The talk page also has a much longer argument, mostly between two editors, about whether or not the Thai name of the site should be included in the text of the article. This debate also recapitiulates issues of which country has the stronger claim for ownership of the temple. Amid accusations of nationalism among the editors, one of them asks "Can we put aside politics and be more collaborative on the encyclopedia project?" and appeals to Wikipedia policies 'Third opinion' and 'Dispute resolution' to diffuse the tensions and refocus attention on the common goal. Beyond these two pages we found that the majority of content on CS-WHL article talk pages is about highly local topics, mostly article administration, editing processes, and quality control.

```{r wh-wiki-map, fig.cap="Wikipedia articles for Cultural Sites on the UNESCO World Heritage List. Countries coloured black currently have no Wikipedia articles for CS-WHL sites. Inset shows the distribution of articles per country. Map data from naturalearthdata.com", fig.width=15}
source(here::here("analysis/paper/002-Wikipedia-CS-WHL-world-map-counts.R"))
knitr::include_graphics(here::here("analysis/figures/wh_country_wikipedia_count_map.png"))
```

@@

Spatial patterns in the coverage of CS-WHL sites on Wikipedia closely resemble the physical distribution of sites. Figure \@ref(fig:wh-wiki-map) shows that most countries have between 2-3 CS-WHL on Wikipedia, with Germany, India, China, France and Spain being the countries with the most articles about their CS-WHL. This pattern is consistent with @graham2014uneven's observations that places in the Global North tend to be over-represented on Wikipedia. More surprising is the distribution of editor locations shown in figure \@ref(fig:wh-wiki-circle), which we determined by geo-locating the IP addresses attached to each anonymous edit on all CS-WHL articles. This figure shows the flow of edits, with the arrows starting at the country where the editor was located when they made their edit, and ending at the country where the CS-WHL site that they are editing is located. The inset plot on figure \@ref(fig:wh-wiki-circle) shows that this visualisation only accounts for a relatively small proportion of all edits to CS-WHL articles. Around 20% of edits of most CS-WHL articles are anonymous, a higher proportion than our random sample. Nevertheless, there are `r prettyNum(nrow(revision_history_page_details_ip_addresses_df), big.mark=",")` anonymous edits, and so this sample likely contains a non-random signal. The most striking detail in figure \@ref(fig:wh-wiki-circle) is the large proportion of edits that originate in the United States, and that the vast majority of these are edits of articles about CS-WHL sites located elsewhere in the world. The country with the next largest proportion, the United Kingdom, is less than half of the United States, and a much greater proportion of edits originating in the UK are on articles about CS-WHL sites located in the UK, compared to the US. The country with the next highest proportion of editors is India, and nearly half of those edits are on CS-WHL sites located India. The majority of countries in this chord diagram indicate that articles about CS-WHL sites in those countries receive edits from editors located elsewhere, mostly the US and other English-speaking countries.

```{r wh-wiki-time1, fig.cap="...", fig.width=15}
source(here::here("analysis/paper/008-Wikipedia-WHL-time-series.R"))
knitr::include_graphics(here::here("analysis/figures/wh_wikipedia_time_inscription.png"))
```

```{r wh-wiki-time2, fig.cap="...", fig.width=15}
knitr::include_graphics(here::here("analysis/figures/wh_wikipedia_time_series_facet.png"))
```

Time series analysis of edits on articles about CS-WHL sites can help us identify relationships between activity on Wikipedia articles and external events related to the sites. Figure \@ref(fig:wh-wiki-time1) shows that most articles about CS-WHL sites are much older than articles in our random sample, with most CS-WHL articles created 5-10 years after Wikipedia first appeared in 2001. This rapid addition of articles early in the life of Wikipedia supports the observation from the in-bound Wikilinks data that articles about CS-WHL sites are more central to the encyclopedia and considered worthy of inclusion than many other types of articles. The distribution of years between the date of a site's inscription on the WHL and the date of the creation of that site's Wikipedia article has a distinctive trimodal shape. The first two modes reflect the relatively large number of sites inscribed on the WHL in 1983 and the early 1990s. The third mode at the zero point on the horizontal axis corresponds to articles that were created in or around the year the CS-WHL site was inscribed on the list. These are sites inscribed after the creation of Wikipedia in 2001. Only a relatively small proportion of CS-WHL sites had Wikipedia articles created about them before they were inscribed on the WHL, because the majority of sites were inscribed before Wikipedia was created. The time series of editing activity at individual sites are highly variable and show display complex patterns (Figure \@ref(fig:wh-wiki-time2)). Timbuktu shows spikes of activity right after attacks in mid June 2012, with descriptions of the attacks added to the text of the article. However the tallest peak in editing activity on the Timbuktu occurs earlier, in April 2012 that  concentrated on expanding . A detailed analysis of these is beyond the scope of this paper, but we can see several patterns that suggestion interesting relationships between editorial activity and WHL listing. For example, editing activity spiking sharply at the time of WHL inscription (e.g. Tyre, Masada), editorial activity spiking at the time of WHL inscription and slowly decaying to a baseline level (e.g. Carthage), editorial activity that peaks shortly after the time of WHL inscription (e.g. Troy, Masada), and many sites where editorial activity that shows no signal at all at the time of WHL inscription (Figure \@ref(fig:wh-wiki-time2)).

@@

Come back to danger list [@brown2019politics; @holleland2019naming]

Sites where the conflict is international have it show on WP

https://en.wikipedia.org/wiki/Wikipedia:Geopolitical,_ethnic,_and_religious_conflicts_noticeboard now defunct

@@

>>>>

---
title: "World Heritage Sites on Wikipedia: Controversies in a Context of Constrained Agency for Cultural Heritage Activism"
author:
  - Ben Marwick
  - Prema Smith
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  UNESCO World Heritage sites are places of outstanding significance, and often key sources of information that influence how people interact with the past today. The process of inscription on the UNESCO list is complicated and intersects with political and commercial controversies. But how well are these controversies known to the public? Wikipedia pages on these sites offer a unique dataset for insights into public understanding of heritage controversies. The unique technicity of Wikipedia, with its bot ecosystem and editing mechanics, shapes how knowledge about cultural heritage is constructed, and how controversies are negotiated and communicated. In this article we investigate the patterns of production, consumption and spatial and temporal distributions of Wikipedia pages for World Heritage cultural sites. We find that Wikipedia provides a distinctive context for investigating how people experience and relate to the past in the present. The agency of participants is highly constrained, but distinctive, behind-the-scenes expressions of cultural heritage activism are evident. Concerns about state-like actors, violence and destruction, deal-making, etc. in the World Heritage inscription process are poorly represented on Wikipedia’s World Heritage pages, instead hyper-local, and process issues dominate controversies on Wikipedia. We will consider how this kind of research, drawing on big data and data science methods, adds new dimensions to the debate on heritage controversies, especially its unique contributions, and also its limitations.
---

@@

An additional consideration for understanding participation in online peer production communities are the technical schemas of the software that Wikipedia runs on, MediaWiki. This is a complex toolkit that enables participation in Wikipedia in highly structured ways. On one hand, these structured behaviours produce structured datasets that are well suited to data science methods for efficient computational analysis of large numbers Wikipedia articles. On the other hand, they constrain and limit the agency of the user, canalizing their behaviour into a small number of possible actions and acceptable modes of discourse and engagement with other users [@iba2010analyzing]. While Wikipedia has elements that are ubiquitous on the Internet, such as links on articles that take the user to other articles or other pages on the Internet, it also has several less common elements that contribute to its unique technicity, resulting in very specific types of relationships between human users and the technical elements of the Wikipedia project [@niederer2010wisdom; @weltevrede2016platform]. For example, every edit to an article is tracked in a publicly accessible version control system attached to that article. This exposes the article creation process in highly granular detail; for any given article, we can see how many editors contributed, the size of their edits and their distribution over time, among other things [@priedhorsky2007creating]. Wikipedia has a special category of edit called the 'revert' which allows a user to restore an article to an earlier state to remove recent vandalism (such as the addition of irrelevant or offensive material). This special revert action, combined with a 'talk' page attached to each encyclopedia article for threaded discussion among editors, allows us to detect and study the dynamics of conflicts arising from the creation and editing of articles [@suh2007us]. For example, a high proportion of 'revert' type edits, and a high word count on the talk page can indicate the controversiality of an article [@yasseri2012dynamics].

@@

In addition to the human users and the technological system that enables and constrains their activities on Wikipedia, there is an important third element of the ecosystem that contributes to Wikipedia's uniqueness: the bots. Wikipedia bots are computer scripts that automatically handle repetitive and mundane tasks to develop, improve, and maintain the encyclopedia. While bots are not unique to Wikipedia, they are important contributors, responsible for a large proportion of edits, and shape the human experience in significant ways [@niederer2010wisdom; @geiger2009social; @geiger2014bots]. They also evolve and autonomously engage in complex interactions with other bots to modify the encyclopedia [@tsvetkova2017even; @geiger2017operationalizing]. @zheng2019roles studied all 1,601 bots that were registered at the time of their study, and classified them into 9 bot roles and 25 associated bot functions. Table \@ref(tab:bot-table) shows the bots that were identified in our study and their categories and functions according to @zheng2019roles.

@@

We obtained data about Wikipedia articles by scraping the HTML pages with the rvest package for R [@rvest]. We used the SelectorGadget [@cantino2017selectorgadget] extension for the Chrome web browser to identify specific elements of interest, or nodes, on the HTML pages, and wrote custom R functions to extract data from these nodes. Our entry points were the Wikipedia articles that are lists of World Heritage regions in major geographical regions of the globe. We found fifteen of these, and scraped the CS-WHL site names from the tables on these pages, and followed the links to scrape the article text, edit history and talk page text for each CS-WHL site included on those tables. A small number of CS-WHL sites have Wikipedia articles that are not included on these tables, we did not include these in our sample. We chose this approach because the Wikipedia article titles for CS-WHL sites often differ from the official site name on the UNESCO list, making manual look-ups very inefficient.

The highly structured nature of most Wikipedia pages afforded us much efficiency during data collection, however, a not insignificant number of pages had minor irregularities, for example, exotic table formatting or minor deviations from widely used templates, that necessitated more complex programming. An especially challenging issue was minor irregularities in spelling place names, for example, "Saint Kitts and Nevis" and "St. Kitts and Nevis", which can confound programmatic joining and summarizing of data by failing to recognize as the same place and leading to under-counting. We attempted to correct these manually in our code. Another challenge was in the formatting of tables, for example a merged cell in a table introduced an irregularity that made parsing and scraping complicated and fragile. We collected data during May 2019, and due to the dynamic nature of Wikipedia, it is highly likely that articles in our study have subtly changed since our data collection as users and bots continue their activity. It is unlikely that articles in our sample have been deleted from Wikipedia since our data collection, because of the high notability of CS-WHL, but it is possible that new articles have since appeared about sites that were not represented on Wikipedia since we collected the data. Our original scraping code may no longer work as the tables on Wikipedia articles continue to be modified.

@@

Only one of the controversial sites discussed above is in the group of articles where the talk page is longer than the article itself, and none of the site on the UNESCO heritage in danger list are in this group. The talk page for Mapungubwe is lengthy with expressions of concern about the accuracy of information in the article. In particular, editors were concerned about the depiction of Indigenous people, especially the degree to which early European colonists were aware of Indigenous communities, and the correct cultural affiliation of the Indigenous groups that originally occupied the site. These issues closely echo the themes identified by @meskell2011paris in her analysis of the WHC process. Although not notable for the length of its talk page, the discussions among editors of the article about Preah Vihear Temple include a strongly worded disagreement among seven editors about the legitimacy of the ownership claims by Thailand and Cambodia of the territory that includes the temple. The debate among Wikipedia editors mirrors very closely the political tensions surrounding the nomination of Preah Vihear. The talk page also has a much longer argument, mostly between two editors, about whether or not the Thai name of the site should be included in the text of the article. This debate also recapitiulates issues of which country has the stronger claim for ownership of the temple. Amid accusations of nationalism among the editors, one of them asks "Can we put aside politics and be more collaborative on the encyclopedia project?" and appeals to Wikipedia policies 'Third opinion' and 'Dispute resolution' to diffuse the tensions and refocus attention on the common goal. Beyond these two pages we found that the majority of content on CS-WHL article talk pages is about highly local topics, mostly article administration, editing processes, and quality control.

```{r wh-wiki-map, fig.cap="Wikipedia articles for Cultural Sites on the UNESCO World Heritage List. Upper panel shows the total number of Wikipedia articles for CS-WHL sites per country. Lower panel is the proportion of all CS-WHL sites that have Wikipedia articles per country. Inset shows the distribution of numbers of CS-WHL articles per country. Map data from naturalearthdata.com", fig.width=15}
source(here::here("analysis/paper/002-Wikipedia-CS-WHL-world-map-counts.R"))
knitr::include_graphics(here::here("analysis/figures/wh_country_wikipedia_count_prop_map.png"))
```

@@

Spatial patterns in the coverage of CS-WHL sites on Wikipedia closely resemble the physical distribution of sites. Figure \@ref(fig:wh-wiki-map) shows that most countries have between 2-3 CS-WHL on Wikipedia, with Germany, India, China, France and Spain being the countries with the most articles about their CS-WHL. This pattern is consistent with @graham2014uneven's observations that places in the Global North tend to be over-represented on Wikipedia. The lower panel of figure \@ref(fig:wh-wiki-map) further demontrates this with South American countries having low proportions of their CS-WHL sites represented on Wikipedia. The pattern in African countries is that they either have a very small number of sites that are all on Wikipedia (indicated by yellow on the figure), or they do not have any CS-WHL at all. More surprising is the distribution of editor locations shown in figure \@ref(fig:wh-wiki-circle), which we determined by geo-locating the IP addresses attached to each anonymous edit on all CS-WHL articles. This figure shows the flow of edits, with the arrows starting at the country where the editor was located when they made their edit, and ending at the country where the CS-WHL site that they are editing is located. The inset plot on figure \@ref(fig:wh-wiki-circle) shows that this visualisation only accounts for a relatively small proportion of all edits to CS-WHL articles. Around 20% of edits of most CS-WHL articles are anonymous, a higher proportion than our random sample. Nevertheless, there are `r prettyNum(nrow(revision_history_page_details_ip_addresses_df), big.mark=",")` anonymous edits, and so this sample likely contains a non-random signal. The most striking detail in figure \@ref(fig:wh-wiki-circle) is the large proportion of edits that originate in the United States, and that the vast majority of these are edits of articles about CS-WHL sites located elsewhere in the world. The country with the next largest proportion, the United Kingdom, is less than half of the United States, and a much greater proportion of edits originating in the UK are on articles about CS-WHL sites located in the UK, compared to the US. The country with the next highest proportion of editors is India, and nearly half of those edits are on CS-WHL sites located India. The majority of countries in this chord diagram indicate that articles about CS-WHL sites in those countries receive edits from editors located elsewhere, mostly the US and other English-speaking countries.

```{r wh-wiki-time1, fig.cap="Article editing relative to WHL inscription date. Upper panel is a histogram of WHL inscription years. Lower panel is a histogram of duration between inscription and the appearance of the Wikipedia article. Inset shows the distribution of CS-WHL article ages and ages of random sites.", fig.width=15}
source(here::here("analysis/paper/008-Wikipedia-WHL-time-series.R"))
knitr::include_graphics(here::here("analysis/figures/wh_wikipedia_time_inscription.png"))
```

```{r wh-wiki-time2, fig.cap="Time series of editing activity for CS-WHL articles for all sites inscribed after Wikipedia began in 2001, and with a minimum of 150 edits in total", fig.width=15}
knitr::include_graphics(here::here("analysis/figures/wh_wikipedia_time_series_facet.png"))
```

Time series analysis of edits on articles about CS-WHL sites can help us identify relationships between activity on Wikipedia articles and external events related to the sites. Figure \@ref(fig:wh-wiki-time1) shows that most articles about CS-WHL sites are much older than articles in our random sample, with most CS-WHL articles created 5-10 years after Wikipedia first appeared in 2001. This rapid addition of articles early in the life of Wikipedia supports the observation from the in-bound Wikilinks data that articles about CS-WHL sites are more central to the encyclopedia and considered worthy of inclusion than many other types of articles. The distribution of years between the date of a site's inscription on the WHL and the date of the creation of that site's Wikipedia article has a distinctive trimodal shape. The first two modes reflect the relatively large number of sites inscribed on the WHL in 1983 and the early 1990s. The third mode at the zero point on the horizontal axis corresponds to articles that were created in or around the year the CS-WHL site was inscribed on the list. These are sites inscribed after the creation of Wikipedia in 2001. Only a relatively small proportion of CS-WHL sites had Wikipedia articles created about them before they were inscribed on the WHL, because the majority of sites were inscribed before Wikipedia was created. The time series of editing activity at individual sites are highly variable and show display complex patterns (Figure \@ref(fig:wh-wiki-time2)). Timbuktu shows spikes of editing activity right after the rebel attacks in mid June 2012, with descriptions of the attacks added to the text of the article at that time. However the tallest peak in editing activity on the Timbuktu occurs earlier, in April 2012, those edits concentrated on expanding the sections on local prehistory. A detailed analysis of all the article-specific time series is beyond the scope of this paper, but we can see several patterns that suggestion interesting relationships between editorial activity and WHL listing. For example, editing activity spiking sharply at the time of WHL inscription (e.g. Tyre, Masada), editorial activity spiking at the time of WHL inscription and slowly decaying to a baseline level (e.g. Carthage), editorial activity that peaks shortly after the time of WHL inscription (e.g. Troy, Masada), and many sites where editorial activity that shows no signal at all at the time of WHL inscription (Figure \@ref(fig:wh-wiki-time2)).

@@

Our results establish Wikipedia as site and community demonstrating active engagement with cultural heritage [@pentzold2017digging], especially cultural sites on the UNESCO World Heritage List. We have defined Wikipedia as an online peer production community with technical and cultural qualities that separate it from social media such as Twitter and Facebook. Editing mechanics, bots, reverts, and talk pages work together to enable and limit discourse and debate about CS-WHL on Wikipedia. The goal-driven culture of Wikipedia editors, and the organisation's policies that guide the production of articles strongly constrain heritage discourse, such that the exchange of different perspectives are rarely found in articles themselves, but buried in technical and highly specific discussions on article talk pages. Our results show how these distinctive qualities generate a technicity that shapes how people engage with elements of the past and attribute social and cultural meanings on Wikipedia. These resemble the "contingent collaborations" and "productive frictions" described by @tsing2011friction where local, individual actions of knowledge creation are circumscribed by forces and structures that encompass their expression, but also give meaning to their interactions.

We found that many of the Wikipedia articles about CS-WHL are central to the encyclopedia project. Many of them were among the first articles written for the encyclopedia, are more frequently viewed, are more frequently linked to by other articles, are longer, and more intensively edited by a greater variety of editors, than the average article. These results indicate the high value that Wikipedia users place on CS-WHL, consistent with the goal of the UNESCO World Heritage Committee to maintain a list reflects the world's cultural diversity of outstanding universal value. However, we found that sites are unequally valued, in predictable ways, given prior work on the demographics of Wikipedia users and editors, and the Eurocentric value system of the World Heritage Committee [@cleere2003uneasy; @smith2006uses p. 98]. CS-WHL in countries in the Global South are under-represented on Wikipedia, a geographical tension that is also evident in the World Heritage List [@brumann2014shifting]. They are also shorter and received less attention in the form of page views and inward Wikilinks.

Perhaps the most striking spatial pattern in our results is how much of the edits to CS-WHL articles in any part of the world come from the United States. The anonymous editors of articles about CS-WHL are generally not located in the same country as the site they are editing. This suggests a digital colonialism of world heritage information on Wikipedia, where heritage is interpreted and communicated by people who are not part of the descendant community traditionally associated with the site. This is an important observation because it is these editors that control the production of knowledge about CS-WHL articles, and the ability to control information about heritage has several implications. First is that the choices of these editors define community identity, that is, the community of Wikipedia editors, and also the broader community of Internet users that read Wikipedia, by indicating which sites are important and visible, and which are not by contributing to a digital 'authorized heritage discourse' [@smith2012constrained] that prioritizes their own self-interests. Second is that traditional communities associated with CS-WHL have little control over how their own heritage is represented on Wikipedia. We see this spatial pattern on Wikipedia as part of a trend of globalization of heritage, where places that had traditionally contributed to local or regional identities are turned into sanitized playgrounds for rich tourists [@bernbeck2004political]. Wikipedia may be seen as virtual playground for Americans to construct a universal heritage that reflects their neo-liberal cultural values, and validates their identities. In this way, the representation of heritage on Wikipedia recapitulates a long-standing tradition of colonial subordination of one group's heritage and identity by a more powerful and dominant group.

This colonial logic may help to explain our findings of minimal conflicts surrounding the production of Wikipedia articles about CS-WHL. We found that bot activity, numbers of revert edits, and edits relating to vandalism for CS-WHL articles were comparable to or less than other sites. Most of this activity relates to playful vandalism and occurred on uncontroversial CS-WHL sites such as the Sydney Opera House and the Tower of London. Controversial sites, identified by the presence of military or diplomatic conflicts, were not prominent in these metrics. Our analysis of the talk pages of CS-WHL articles shows that they are often longer than a typical Wikipedia article, and are the location where conflict is most evidence. While much of the conflict is highly technical and local about the policies and processes of editing Wikipedia, we did observe signs of the political and military conflict at CS-WHL played out on some of the talk pages, such as Mapungubwe and Preah Vihear. We interpret this as editors exploiting the talk pages for behind-the-scenes heritage activism. The technical and cultural constraints of writing directly on the article page mean that the talk page is the only suitable location for negotiating disputes whose eventual trace in the edit history of the article might be negligible. We found that some CS-WHL talk pages referenced a 'Geopolitical, ethnic, and religious conflicts noticeboard' elsewhere on Wikipedia, as a forum for arbitration of disputes created in 2008 following a recommendation by an investigation into edit wars on articles about Northern Ireland (The Troubles), Palestine-Israel and other locations with ethnic or cultural based conflicts. However the noticeboard was closed in 2012 due to an absence of activity. Overall we find that editors of CS-WHL articles appear conflict-averse, probably because many of them come from similar cultural, socio-economic and educational backgrounds, and so share similar values about heritage that align with their shared goal of making an encyclopedia. The low levels of conflict reflect the continuity between the printed encyclopedia project as a compendium of universal knowledge, and the World Heritage List as universal heritage. These projects share origins in the cultural elites of Western European, and their distinctive sense of what is universal that is linked to nineteenth-century nationalism and liberal modernity [@smith2006uses].

The results of our time series analyses reveal complex patterns in how heritage-related concerns are realized on Wikipedia. Some sites show spikes in Wikipedia editing activity that are directly related to the inscription of the site on the WHL, or conflict events at the site. However at other sites the rhythms of editing activity seem unrelated to events directly related to the site. Like the talk pages, which have unrealised potential for close reading of heritage activism and disputes, close study individual edit histories of CS-WHL articles may reveal further patterns of how people engage with heritage on Wikipedia.

@@

The agency of people participating in heritage-related activities on Wikipedia is highly constrained by the technical characteristics of editing, bots, and consensus-valuing culture. Although there are some niches where editors can debate, concerns about state-like actors, violence & destruction, deal-making, etc. are poorly represented. Typical conflicts about Wikipedia CS-WHL articles are hyper-local, and process-centered. Spatial patterns in the edits to Wikipedia articles about CS-WHL show a kind of digital colonialism of world heritage information, with most edits emanating from the United States. Because of these technical and cultural characteristics of Wikipedia, its articles present a discourse that actively obscures the power relations that give rise to them, and makes opaque debates about how heritage is involved in the production of identity and power [cf.@harvey2001heritage].

That said, Wikipedia is valuable and important as a Big Data source of social information on heritage for several reasons. Although it lacks the velocity of other social media, such as Twitter, it preserves a distinctively complex record of human interactions with the past at simultaneously high scales, high resolution, and in highly structured ways, afforded by no other platform. As a non-profit entity, Wikipedia's data are freely available to researchers, unlike Twitter and Facebook. This has important ethical implications because it eliminates the need for academic collaboration with commercial platforms, whose business interests may not align with the researchers' ethics [@schroeder2014big]. As a context of heritage production, a resource of information on heritage, and an artefact of contemporary digital heritage, Wikipedia is remarkable and holds substantial research potential. Given the tremendous popularity of Wikipedia, is it important that future work on digital heritage in this context continue to critically examine what is absent from, and uncontested in, Wikipedia, and who benefits and suffers from these absences.
